A Compound Decision Approach to Covariance Matrix Estimation

### 2022-03-07 msg.tex
change of to for
biometrics second revision, updated to satisfy biometrics editor comments only

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications. However, in modern problems the number of features can be of the same order as or exceed the sample size. This high-dimensional setting is especially common in genomics, where covariance matrices are used to model gene networks but the number of genes can be much larger than the number of biological replicates \citep{schafer2005shrinkage, markowetz2007inferring}. For example, in Section \ref{gene analysis} we study brain region-specific gene networks in mouse using bulk RNA-sequencing data.

A common approach to estimating gene expression networks is to use the standard sample covariance or corrlation matrices \citep{langfelder2008wgcna, zhang2005general}. However, these estimators behave poorly in the high-dimensional regime. To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrices. These can roughly be divided into two groups, according to whether they impose assumptions of the structure of population covariance matrix.

@@

In this paper we define and explore one such class of estimators. This approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}. Section \ref{sec:compound} shows how this framing motivates vectorizing the covariance matrix and solving the resulting vector estimation problem using nonparametric empirical Bayes procedures studied in the compound decision literature \citep{jiang2009general, koenker2014convex, efron2019bayes}. The vector estimate can then be reassembled into matrix form and projected into the space of positive-definite matrices to give a final covariance matrix estimator. Surprisingly, though this vectorized approach essentially ignores the matrix structure, it can sometimes outperform a number of state-of-the-art proposals in simulations and a real data analysis. Our results suggest that vectorized approaches may be preferrable when the sample eigenvectors are poor estimates of the population eigenvectors. Our compound decision approach is closely related to the correlation matrix estimator of \citet{dey2018corshrink}. We discuss in detail the connections with their work in Section \ref{sec:proposed}.

The article is organized as follows. In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed estimator. In Section \ref{numerical results} we illustrate its performance in simulations and a gene coexpression network inference problem. Finally, Section \ref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub at \url{http://github.com/sdzhao/cole}.

@@

Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$. This section argues that covariance matrix estimation can be viewed as a compound decision problem when estimation performance is measured using the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate. This compound decision formulation will motivate a new class of unstructured non-rotationally invariant estimators.

Compound decision problems involve the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$ \citep{robbins1951asymptotically}. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk $ n^{-1} \sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))$, where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the classical Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.

Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can therefore be viewed as a compound decision problem where the goal is to simultaneously estimate the components of the vector
\begin{equation}
  \label{eq:sigma}
  \bs{\sigma} = (\sigma_{11}, \ldots, \sigma_{1p}, \ldots, \sigma_{p1}, \ldots,\sigma_{pp})^\top
\end{equation}
under average squared error loss.  One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite. Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}. As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite. This will in fact be true of one of the estimators we propose in Section \ref{sec:proposed}. To resolve this issue, we can project any estimate into the space of positive-definite matrices; see Section \ref{posdef}.

The value of this formulation is that it allows us to apply ideas from the compound decision literature to covariance matrix estimation.  A key property of compound decision problems is that while information about index $i$ seems like it should offer no help for estimating a parameter with a different index $j$, in fact borrowing information across all indices is superior to considering each index alone. An example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. \citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$. A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}. A natural next step is therefore to apply modern ideas for Gaussian sequence estimation to covariance matrix estimation, which we do in Section \ref{sec:method}.

@@

Treating the covariance matrix estimation problem as a vector estimation compound decision problem may seem unintuitive, but many existing matrix estimation methods can also be reinterpreted as carrying out vector estimation. The simplest example is the sample covariance matrix $\bs{S} = n^{-1} \bs{X}^{T}\bs{X}$, which takes this form because the $\bs{X}$ are assumed to have mean zero. As another example, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective. Their adaptive thresholding method applies a version of the soft thresholding method of \citet{donoho1995adapting}, originally developed to estimate a sparse mean vector in the Gaussian sequence problem, to each entry of the sample covariance matrix.

As a further illustration of the potential of our new formulation, we can show that the linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as the solution to a compound decision problem. As described in Section \ref{sec:background}, we first view the problem of minimizing the scaled Frobenius norm \eqref{frobenius risk} as simultaneously estimating the components of the vector $\bs{\sigma}$ \eqref{eq:sigma} under squared error. Then a standard approach in the compound decision literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes the compound risk. For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule \citep{fourdrinier2018shrinkage, stigler19901988}.

Following this recipe, we also restrict attention to the class of linear decision rules where $\sigma_{jk}$ will be estimated by
\begin{equation}
  \label{linear class}
  \delta_{jk}(\bs{X}) = \beta_S s_{jk} + \beta_I u_{jk}
\end{equation}
for some parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of $\bs{I}$. Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}. Though we cannot directly calculate \eqref{frobenius risk} because it involves the unknown $\sigma_{jk}$, the following proposition shows that we can construct a good estimate of this risk for every fixed value of $\beta_S$ and $\beta_I$.

\begin{prop}
  \label{prop:Rhat}
  Define $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ and
  \begin{equation}
    \label{eq:Rhat}
    \hat{R}_L(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ].
  \end{equation}
  If $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$, then
  \[
    \lim_{n \rightarrow 0} \{ \bb{E} \hat{R}_L(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta}) \} = 0
  \]
  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}. $\Vert \cdot \Vert_F$ denotes the Frobenius norm $\Vert \bs{A} \Vert_F^2=\sum_{j,k=1}^p A_{jk}^2$.
\end{prop}
The condition on $\bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ requires that the variances of the entries of $\bs{S}$ do not grow too quickly as $n$ grows. It is also required by \citet{ledoit2004well} and is implied by their Lemma 3.1.

Proposition \ref{prop:Rhat} shows that $\hat{R}_L(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}. It is then reasonable to estimate the best estimator in this class by minimizing $\hat{R}_L(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. A slight modification of this procedure turns out to recover the \citet{ledoit2004well} estimator
\begin{equation}
\label{eq:lw}
\hat{\bs{\Sigma}}_{\text{LW}} = \left(1 - \frac{b_n^2}{d_n^2}\right) \bs{S} + \frac{b_n^2}{d_n^2}\hat{\mu} \bs{I},
\end{equation} 
where $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$.

@@

There are also interesting connections between \eqref{eq:opt_linear} and classical estimators of a vector of normal means. First, instead of \eqref{eq:opt_linear}, consider the estimator $\tilde{\delta}_{jk}(\bs{X}) = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, which directly plugs the estimated $\hat{\beta}_S$ and $\hat{\beta}_I$ into \eqref{linear class}. It can be shown that  $\tilde{\delta}_{jk}(\bs{X})$ is equivalent to estimating $\bs{\sigma}$ by shrinking the vector $(s_{11}, \ldots, s_{1p}, \ldots, s_{p1}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots,  u_{1p}, \ldots, u_{p1}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion, stigler19901988}. A major difference between $\hat{\delta}_{jk}$ \eqref{eq:opt_linear} and $\tilde{\delta}_{jk}$ is that $\hat{\beta}_S$ in the latter is replaced by $\max(\hat{\beta}_S, 0)$ in the former. This is similar to how replacing the shrinkage factor in the classical James-Stein estimator by its positive part results can improve its accuracy \citep{baranchik1964multiple}.

@@

Proposition \ref{prop:linear} of Section \ref{sec:connections} suggests that applying ideas from the compound decision literature to estimating the vector $\bs{\sigma}$ \eqref{eq:sigma}  can be a fruitful strategy for covariance matrix estimation. However, Proposition \ref{prop:linear} considers only a linear decision rule, whereas more recent work in compound decision theory has focused on the much larger class of so-called separable rules \citep{brown2009nonparametric, jiang2009general, zhang2003compound}. In the standard compound decision problem of estimating $\bs{\theta} = E \bs{Y}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ for some function $t$ that does not depend on the index $i$ \citep{robbins1951asymptotically}. Linear rules such as \eqref{eq:opt_linear} are included in this class, so that the best separable rule should perform at least as well as the best linear rule.

Motivated by these ideas, here we propose a new class of covariance matrix estimators by generalizing the notion of separable rules to the problem of estimating a vectorized matrix. For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, we define the class of separable rules to be
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
   \delta_{jj} = t_d(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
\end{equation}
where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature. In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $t_d$ and off-diagonal entries using $t_{od}$, where $t_{od}$ and $t_d$ do not depend on the indices $j$ and $k$. Furthermore, we enforce rule in $\mathcal{S}$ to give symmetric estimates of the off-diagonal entries.

@@

This class of separable rules is reasonable to consider because it includes several common covariance estimators, including the sample covariance, the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}, and the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
\begin{align*}
t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}) &= \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n, \, 1\leq k<j\leq p,\\
t_d(\bs{X}_{\cdot j})&=\beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot j} / n+\beta_I, \, j=1,\ldots,p.
\end{align*}
Furthermore, whereas these three existing estimators are ultimately all functions of the $s_{jk}$, the class $\mathcal{S}$ allows for much more general rules that can be any function of the observations $\bs{X}_{\cdot j}$ and $\bs{X}_{\cdot k}$.

The class $\mathcal{S}$ constitutes a fundamentally different approach to covariance matrix estimation compared to the existing methods described in Section \ref{introduction}. Estimators in $\mathcal{S}$ do not assume that $\bs{\Sigma}$ has any particular structure and also are not necessarily rotationally invariant. Intuitively, they should perform better than rotationally invariant estimators when the sample eigenvectors of $\bs{S}$ are very different from the population eigenvectors. Results in Section \ref{numerical results} suggest that this is indeed the case.

@@

Instead, we use a clustering-based exemplar algorithm to further improve computational efficiency. Let $s_j$, $s_k$, and $\gamma_{jk}$ be the sample standard deviations and correlation between the $j$th and $k$th covariates. We first apply $K$-means clustering to identify $K$ cluster centroids using the $p (p-1) / 2$ lower triangular off-diagonal sample points $(s_j,s_k,\gamma_{jk})$. We then swap the first two coordinates of each point to construct another set of $K$ centroids. Finally, we use these $2K$ centroids as the support points for $\hat{G}_{od}$. We evaluate the impact of changing $K$ in Section \ref{numerical results}.

Our procedure is implemented in the R package \verb|cole|, available on GitHub. In our implementations of the EM algorithm in Section \ref{numerical results} we stopped iterating when the relative difference between two successive log-likelihoods was less than $10^{-4}$, up to a maximum of 200 iterations. Instead of EM, we attempted to maximize the composite likelihood \eqref{Gnd hat} using the interior point solver MOSEK, as advocated by \citet{koenker2014convex}, but we ran into computational difficulties because for even moderate sample sizes $n$, the values of the densities $f_{1n}$ and $f_{2n}$ at certain support points could be extremely small. These small values caused problems for the R wrapper for MOSEK. Our implementation of the EM algorithm was more robust to these small density values.

@@

\begin{itemize}
\item Model 1 (sparse): we generated a sparse $\bs{\Sigma}$ where most features were not correlated. We let half of the standard deviations equal 1 and the other half equal 1.5. We set the correlation matrix following Model 2 of \citet{cai2011adaptive}:
  \[
  \begin{pmatrix}&\bs{A}_1 & \bs{0} \\ &\bs{0} &\bs{I}_{p/2\times p/2}\end{pmatrix},
  \]
  where the $jk$th entry of $\bs{A}_1$ is $a_{jk} = \max(1- \vert j - k \vert / 10, 0)$.

\item Model 2 (hypercorrelated): we generated a $\bs{\Sigma}$ that we call ``hypercorrelated'' because its parameters are in a sense dependent on each other, as we let larger standard deviations $\sigma_j$ and $\sigma_k$ correspond to larger $r_{jk}$. Specifically, we let the first $p / 2$ standard deviations equal 1, the last $p / 2$ equal 2, and the correlation matrix equal
  \[
    \begin{pmatrix} & \bs{C}_{11} & \bs{C}_{12} \\  & \bs{C}_{21} & \bs{C}_{22}\end{pmatrix},
  \]
  where $\bs{C}_{11}$ and $\bs{C}_{22}$ were $p/2 \times p/2$ compound symmetric matrices with correlation parameters 0.8 and 0.2, respectively, and $\bs{C}_{12}$ and $\bs{C}_{21}$ were $p/2 \times p/2$ matrices with entries equal to 0.4. As discussed in Section \ref{sec:proposed}, we expect that our proposed approach will be able to take advantage of the dependencies between the $\sigma_j$ and the $r_{jk}$, unlike the empirical Bayes approach of \citet{dey2018corshrink} that does not shrink standard deviations.

>>>>

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications. However, in modern problems the number of features can be of the same order as or exceed the sample size. This high-dimensional setting is especially common in genomics, where covariance matrices are used to model gene networks but the number of genes can be much larger than the number of biological replicates \citep{schafer2005shrinkage, markowetz2007inferring}. For example, in Section \ref{gene analysis} we study brain region-specific gene networks in mouse using bulk RNA-sequencing data. A common approach to estimating gene expression networks is to use the standard sample covariance or corrlation matrices \citep{langfelder2008wgcna, zhang2005general}. However, these estimators behave poorly in the high-dimensional regime, and there are roughly two types of approaches to address this issue.

@@

In this paper we define and explore one such class of estimators. This approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}. Section \ref{sec:compound} shows how this framing motivates vectorizing the covariance matrix and solving the resulting vector estimation problem using nonparametric empirical Bayes procedures studied in the compound decision literature. The vector estimate can then be reassembled into matrix form and projected into the space of positive-definite matrices to give a final estimate. Surprisingly, though this vectorized approach essentially ignores the matrix structure, it can sometimes outperform a number of state-of-the-art proposals in simulations and a real data analysis. Our compound decision approach is closely related to the correlation matrix estimator of \citet{dey2018corshrink}. We discuss in detail the connections with their work in Section \ref{sec:proposed}.

The article is organized as follows. In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed estimator. In Section \ref{numerical results} we illustrate its performance in simulations and a gene coexpression network inference problem. Section \ref{sec:discussion} concludes with a discussion, and proofs can be found in the Supporting Information.

@@

Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$. A common measure of estimation performance is the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.

Minimizing \eqref{frobenius risk} can be viewed as a compound decision problem. These involve the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$ \citep{robbins1951asymptotically}. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk $ n^{-1} \sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))$, where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the classical Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.

Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} is thus a compound decision problem where the goal is to simultaneously estimate the components of
\begin{equation}
  \label{eq:sigma}
  \bs{\sigma} = (\sigma_{11}, \ldots, \sigma_{1p}, \ldots, \sigma_{p1}, \ldots,\sigma_{pp})^\top
\end{equation}
under average squared error loss.  One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite. Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}. As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite. To resolve this issue, we can project any estimate into the space of positive-definite matrices; see Section \ref{posdef}.

The value of this formulation is that it allows us to apply ideas from the compound decision literature to covariance matrix estimation.  A key property of compound decision problems is that while information about index $i$ seems like it should offer no help for estimating a parameter with a different index $j$, in fact borrowing information across all indices is superior to considering each index alone. An example of this phenomenon is the James-Stein shrinkage estimator \citep{james1961estimation}% , which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. \citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$
, and a long line of subsequent work has led to much more sophisticated and accurate procedures \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}. In Section \ref{sec:method} we apply these ideas to covariance matrix estimation,

@@

Treating covairance matrix estimation as a vector estimation problem may seem unintuitive, but many existing matrix estimation methods can also be reinterpreted as carrying out vector estimation. The simplest example is the sample covariance matrix $\bs{S} = n^{-1} \bs{X}^{T}\bs{X}$, which takes this form because the $\bs{X}$ are assumed to have mean zero. As another example, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective. Their adaptive thresholding method applies a version of the soft thresholding method of \citet{donoho1995adapting}, originally developed to estimate the mean of a Gaussian sequence, to each entry of the sample covariance matrix.

As a further illustration of the potential of our new formulation, we can show that the linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as the solution to a compound decision problem. A standard approach in this literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes the compound risk % . For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule 
\citep{fourdrinier2018shrinkage, stigler19901988}. Following this recipe, we consider the class of linear decision rules where $\sigma_{jk}$ will be estimated by
\begin{equation}
  \label{linear class}
  \delta_{jk}(\bs{X}) = \beta_S s_{jk} + \beta_I u_{jk}
\end{equation}
for some parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of $\bs{I}$. Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}. Though we cannot directly calculate \eqref{frobenius risk} because it involves the unknown $\sigma_{jk}$, the following proposition shows that we can construct a good estimate of this risk for every fixed value of $\beta_S$ and $\beta_I$.

\begin{prop}
  \label{prop:Rhat}
  Define $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ and
  \begin{equation}
    \label{eq:Rhat}
    \hat{R}_L(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ].
  \end{equation}
  If $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$, then
  \[
    \lim_{n \rightarrow 0} \{ \bb{E} \hat{R}_L(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta}) \} = 0
  \]
  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}, where $\Vert \bs{A} \Vert_F$ denotes the Frobenius norm $\bs{A}$.
\end{prop}

The condition on $\bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ requires that the variances of the entries of $\bs{S}$ do not grow too quickly as $n$ grows. It is also required by \citet{ledoit2004well} and is implied by their Lemma 3.1. Proposition \ref{prop:Rhat} shows that $\hat{R}_L(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}. It is then reasonable to estimate the best estimator in this class by minimizing $\hat{R}_L(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. A slight modification of this procedure turns out to recover the \citet{ledoit2004well} estimator
\begin{equation}
\label{eq:lw}
\hat{\bs{\Sigma}}_{\text{LW}} = \left(1 - \frac{b_n^2}{d_n^2}\right) \bs{S} + \frac{b_n^2}{d_n^2}\hat{\mu} \bs{I},
\end{equation} 
where $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$.

@@

% There are also interesting connections between \eqref{eq:opt_linear} and classical estimators of a vector of normal means. First, instead of \eqref{eq:opt_linear}, consider the estimator $\tilde{\delta}_{jk}(\bs{X}) = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, which directly plugs the estimated $\hat{\beta}_S$ and $\hat{\beta}_I$ into \eqref{linear class}. It can be shown that  $\tilde{\delta}_{jk}(\bs{X})$ is equivalent to estimating $\bs{\sigma}$ by shrinking the vector $(s_{11}, \ldots, s_{1p}, \ldots, s_{p1}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots,  u_{1p}, \ldots, u_{p1}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion, stigler19901988}. A major difference between $\hat{\delta}_{jk}$ \eqref{eq:opt_linear} and $\tilde{\delta}_{jk}$ is that $\hat{\beta}_S$ in the latter is replaced by $\max(\hat{\beta}_S, 0)$ in the former. This is similar to how replacing the shrinkage factor in the classical James-Stein estimator by its positive part results can improve its accuracy \citep{baranchik1964multiple}.

@@

Proposition \ref{prop:linear} of Section \ref{sec:connections} suggests that applying ideas from the compound decision literature to estimating the vector $\bs{\sigma}$ \eqref{eq:sigma}  can be a fruitful strategy for covariance matrix estimation. However, Proposition \ref{prop:linear} considers only a linear decision rule, whereas more recent work in compound decision theory has focused on the much larger class of so-called separable rules \citep{brown2009nonparametric, jiang2009general, zhang2003compound}. % In the standard compound decision problem of estimating $\bs{\theta} = E \bs{Y}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ for some function $t$ that does not depend on the index $i$ \citep{robbins1951asymptotically}. Linear rules such as \eqref{eq:opt_linear} are included in this class, so that the best separable rule should perform at least as well as the best linear rule.

We generalize these ideas to propose a new class of covariance matrix estimators. For rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, we define the class of separable estimators to be
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} :
  \quad
  \delta_{kj} = \delta_{jk} = t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p;
  \quad 
   \delta_{jj} = t_d(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
\end{equation}
where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature. In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $t_d$ and off-diagonal entries using $t_{od}$, where $t_{od}$ and $t_d$ do not depend on the indices $j$ and $k$. Furthermore, we enforce rules in $\mathcal{S}$ to give symmetric estimates of the off-diagonal entries.

@@

This class of separable rules is reasonable to consider because it includes several common covariance estimators, including the sample covariance, the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}, and the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
\begin{align*}
t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}) &= \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n, \, 1\leq k<j\leq p,\\
t_d(\bs{X}_{\cdot j})&=\beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot j} / n+\beta_I, \, j=1,\ldots,p.
\end{align*}
Furthermore, whereas these three existing estimators are ultimately all functions of the $s_{jk}$, the class $\mathcal{S}$ allows for much more general rules that can be any function of the $\bs{X}_{\cdot j}$ and $\bs{X}_{\cdot k}$. Finally, the class $\mathcal{S}$ constitutes a fundamentally different approach to covariance matrix estimation compared to the existing methods described in Section \ref{introduction}. Estimators in $\mathcal{S}$ do not assume that $\bs{\Sigma}$ has any particular structure and also are not necessarily rotationally invariant.

 % Intuitively, they should perform better than rotationally invariant estimators when the sample eigenvectors of $\bs{S}$ are very different from the population eigenvectors. Results in Section \ref{numerical results} suggest that this is indeed the case.

@@

Instead, we use a clustering-based exemplar algorithm to further improve computational efficiency. Let $s_j$, $s_k$, and $\gamma_{jk}$ be the sample standard deviations and correlation between the $j$th and $k$th covariates. We first apply $K$-means clustering to identify $K$ cluster centroids using the $p (p-1) / 2$ lower triangular off-diagonal sample points $(s_j,s_k,\gamma_{jk})$. We then swap the first two coordinates of each point to construct another set of $K$ centroids and use these $2K$ centroids as the support points for $\hat{G}_{od}$. Section \ref{numerical results} evaluates the impact of changing $K$.

In our implementations of the EM algorithm in Section \ref{numerical results} we stopped iterating when the relative difference between two successive log-likelihoods was less than $10^{-4}$, up to a maximum of 200 iterations. Instead of EM, we attempted to maximize the composite likelihood \eqref{Gnd hat} using the interior point solver MOSEK, as advocated by \citet{koenker2014convex}, but we ran into computational difficulties because for even moderate sample sizes $n$, the values of the densities $f_{1n}$ and $f_{2n}$ at certain support points could be extremely small. These small values caused problems for the R wrapper for MOSEK. Our implementation of the EM algorithm was more robust to these small density values.

@@

Model 1 (sparse): we generated a sparse $\bs{\Sigma}$ where most features were not correlated. We let half of the standard deviations equal 1 and the other half equal 1.5. We set the correlation matrix following Model 2 of \citet{cai2011adaptive}, a block-diagonal matrix where the $jk$th entry of the first $p/2 \times p/2$ block was $\max(1- \vert j - k \vert / 10, 0)$ and the second $p/2 \times p/2$ block was an identity matrix.

Model 2 (hypercorrelated): we generated a $\bs{\Sigma}$ that we call ``hypercorrelated'' because its parameters are in a sense dependent on each other, as we let larger standard deviations $\sigma_j$ and $\sigma_k$ correspond to larger $r_{jk}$. Specifically, we let the first $p / 2$ standard deviations equal 1, the last $p / 2$ equal 2, and the correlation matrix have four $p/2 \times p/2$ blocks. The diagonal blocks were compound symmetric matrices with correlation parameters 0.8 and 0.2 and the off-diagonal blocks had entries equal to 0.4. As discussed in Section \ref{sec:proposed}, we expect that our proposed approach will be able to take advantage of the dependencies between the $\sigma_j$ and the $r_{jk}$, unlike the empirical Bayes approach of \citet{dey2018corshrink} that does not shrink standard deviations.

Models 3 and 4 (dense-0.7 and dense-0.9): we generated a $\bs{\Sigma}$ where the correlation between all features were 0.7 and 0.9, in contrast to our sparse Model 1. Otherwise, we generated the standard deviations as in Model 1.
